FHJ: A Formal Model for Hierarchical Dispatching and Overriding

### 2017-12-08 NewVersion/response.txt

We thank the reviewers for the valuable comments. We try to address the reviews one by one.

Reviewer A.
- generalize the method overriding semantics
We think this suggestion is useful and valueble, we didn't think about this before. It does produce a
simpler method overriding semantics and definitely can be a future work.

- "renaming can break the subtyping relation"
Naturally renaming can break the subtyping relation, of course, there might be techniques as you said 
that tracks the rename mapping and keeps the subtyping relation, but that's the special case and we 
didn't discuss it in the paper.

- "you do not mention the more direct way in which C++ allows disambiguating methods by qualification..."
In Section 2.1 we have mentioned the C++ qualified name approach, see Sec 2.1 -> Potential Fixes -> 
Static dispatch.

- "C++ allows both overriding and shadowing ..."
Since FHJ does not support overloading, it's model is much simpler than C++, so there's no concern 
of distinguishing overriding and shadowing of inherited methods. But we agree that it could be interesting
to add this feature.

- "the notion of 'original' method...."
Sorry that we didn't make it clear. The term "original method" is supposed to be relative to
"hierarchical overriding method". It does include overrides and maybe a better name (for example
"own method"?) is needed.

- "abusing override self syntax for non-(hierarchical-)override...can be empty?..."
Yes, `I { m() }` is a syntactic sugar for `I { m() override I ... }`. In a real language, this should be 
supported. But in formlization, for simplicity, we union them into one.

- Footnote 4: “new I()()” — macro fail?
For footnote 4, the version submitted is "In Java the corresponding syntax is new I(){}." Maybe the 
version you have is not updated?

- "Figure 6: The rule C-StaticType renders the reduction relation non-deterministic"
Without (C-FReduce), (C-StaticType) IS infinite. But in (C-FReduce), the condition `e \neq new J()`
prevent the infinite loop.

Reviewer B.
- "Eiffel seems to allow renaming ..."
In Eiffel, programmers still have to rename methods to resolve the conflicts, so I don't think it is a 
different approach and actually you can't keep the method with the SAME name.

> MiddleMan approach, "whether this problem is really worth a language extension..."
The example in "The C++ programming language" and stackoverflow are both similar to the MiddleMan
approach (they have the same spirit, just MiddleMan approach uses parameter types to distinguish
different methods, while the others use separate sets of methods to distinguish).
We agree that this name conflict problem is not quite popular in practice. That's also the reason
why we do not have a big case study in the paper, because we didn't find a senario that
unintentional method conflicts are frequently used in one project.   But once this issue is
encountered, the effect could be severe, and that existing workarounds might not help a lot. And
since the C++ books, the trait paper as well as other papers mentioned this issue, it might be a
sign that this problem is still an interesting issue in certain cases, which worth exploring and
taking into consideration.  Finally our work shows that this feature is lightweight, it does not
complicate the semantics a lot and can be easily added to language models.

- "It would strengthen your paper if you would provide an implementation..."
As mentioned in Section 1, we have a prototype implementation available here: 
https://github.com/YanlinWang/MIM/tree/master/Calculus

- "What has been removed from FJ: just downcasts?"
Also constructors and fields.

- "5.2: static invocation confusing..."
* "the method m in J_0 that hierarchically overrides J_1" means, the method m defined like this:
interface J_0 { K m() override J_1 { ... } }
* "e1.J_0@J_1 :: m(e2)" is different from "((J1) e1.m(e2)". The first is a static call while the
second dynamic. The result would be different when the method m gets overridden in an heir of J_0.
* For the question "whether it should be forbidden", we think it is a design choice. If the user
makes  a normal e.m() call, for sure it wouldn't access the super methods. But the static invocation
here gives users the freedom to directly call super methods.

- "p.22: "It breaks structural subtyping.."

Review C.
- "The proposal seems unnecessarily complex. In NextGen and MixGen ..."

- "what do you mean by 'reduction model similar to FJ'?"
Sorry for the English, what we mean is that the reduction steps follow the reduction rules in FJ-like
languages, where no static types are tracked.

- "you mention for the first time 'origin'..."
On page 8, "origin" means where the method is defined, we avoid using "original methods" because it
it not introduced yet on page 8. "original method" is a relative concept to hierarchical overriding 
methods.

- "Section 3.2: subtyping is ANTIsymmetric."
Yes, it is not symmetric, we will delete it.

### 2017-12-08 NewVersion/response.txt

- "What has been removed from FJ: just downcasts?"
Also constructors and fields.

- "5.2: static invocation confusing..."
* "the method m in J_0 that hierarchically overrides J_1" means, the method m defined like this:
interface J_0 { K m() override J_1 { ... } }
* "e1.J_0@J_1 :: m(e2)" is different from "((J1) e1.m(e2)". The first is a static call while the
second dynamic. The result would be different when the method m gets overridden in an heir of J_0.
* For the question "whether it should be forbidden", we think it is a design choice. If the user
makes  a normal e.m() call, for sure it wouldn't access the super methods. But the static invocation
here gives users the freedom to directly call super methods.

- "p.22: "It breaks structural subtyping.."

Review C.
- "The proposal seems unnecessarily complex. In NextGen and MixGen ..." 
Thanks for the references, we need to add them. The models in these references also show that the
static types can be utilized for method dispatch, which is similar to C++. But the complexity of our
model mainly comes from the feature of hierarchical overriding, which is not introduced by previous
papers. We believe that such complexity can simplify the client code written by programmers.

### 2017-12-08 NewVersion/response.txt

We thank the reviewers for the valuable comments. We try to address the reviews one by one.

Reviewer A.
- generalize the method overriding semantics

We think this suggestion is useful and valuable, we didn't think about this before. It does produce a
simpler method overriding semantics and definitely can be a future work.

- "renaming can break the subtyping relation"

We are curious to know if the reviewer has a reference to some work
that tracks the name remapping to preserve the subtyping relation.
It is not obvious to us that this would work smoothly. A reference
would be highly appreciatted.

@@

- "you do not mention the more direct way in which C++ allows disambiguating methods by qualification..."
In Section 2.1 we have mentioned the C++ qualified name approach, see Sec 2.1 -> Potential Fixes -> 
Static dispatch.

- "C++ allows both overriding and shadowing ..."
Since FHJ does not support overloading, it's model is simpler than C++, so there's no concern 
of distinguishing overriding and shadowing of inherited methods. But we agree that it could be interesting
to add this feature.

@@

- "abusing override self syntax for non-(hierarchical-)override...can be empty?..."
Yes, `I { m() }` is a syntactic sugar for `I { m() override I ... }`. In a real language, this should be 
supported. But in formlization, for simplicity, we union them into one.

- "Figure 6: The rule C-StaticType renders the reduction relation non-deterministic"
Without (C-FReduce), (C-StaticType) IS infinite. But in (C-FReduce), the condition `e \neq new J()`
prevents the infinite loop.

@@

Reviewer B.
- "Eiffel seems to allow renaming ..."
In Eiffel, programmers still have to rename methods to resolve the conflicts, so I don't think it is a 
different approach and actually you can't keep the method with the SAME name.

> MiddleMan approach, "whether this problem is really worth a language extension..."
The example in "The C++ programming language" and stackoverflow are both similar to the MiddleMan
approach (they have the same spirit, just MiddleMan approach uses parameter types to distinguish
different methods, while the others use separate sets of methods to distinguish).
We agree that this name conflict problem is not quite popular in practice. But once this issue is
encountered, the effect could be severe, and that existing workarounds might not help a lot. And
since the C++ books, the trait paper as well as other papers mentioned this issue, it might be a
sign that this problem is still an interesting issue in certain cases,
which is worth exploring and taking into consideration.  Finally our work shows that this feature is lightweight, it does not
complicate the semantics a lot and can be easily added to language models.

### 2017-12-19 NewVersion/response.txt

We think this suggestion is useful and valuable, we didn't think about this before. It does produce a
simpler method overriding semantics and definitely can be a future work.

- "renaming can break the subtyping relation"

We are curious to know if the reviewer has a reference to some work
that tracks the name remapping to preserve the subtyping relation.
It is not obvious to us that this would work smoothly. A reference
would be highly appreciatted.

- "you do not mention the more direct way in which C++ allows disambiguating methods by qualification..."

In Section 2.1 we have mentioned the C++ qualified name approach, see Sec 2.1 -> Potential Fixes -> 
Static dispatch.

- "C++ allows both overriding and shadowing ..."

Since FHJ does not support overloading, it's model is simpler than C++, so there's no concern 
of distinguishing overriding and shadowing of inherited methods. But we agree that it could be interesting
to add this feature.

- "abusing override self syntax for non-(hierarchical-)override...can be empty?..."

Yes, `I { m() }` is a syntactic sugar for `I { m() override I ... }`. In a real language, this should be 
supported. But in formlization, for simplicity, we union them into one.

- "Figure 6: The rule C-StaticType renders the reduction relation non-deterministic"

Without (C-FReduce), (C-StaticType) IS infinite. But in (C-FReduce), the condition `e \neq new J()`
prevents the infinite loop.

Reviewer B.

- "Eiffel seems to allow renaming ..."

In Eiffel, programmers still have to rename methods to resolve the conflicts, so I don't think it is a 
different approach and actually you can't keep the method with the SAME name.

- MiddleMan approach, "whether this problem is really worth a language extension..."

The example in "The C++ programming language" and stackoverflow are both similar to the MiddleMan
approach (they have the same spirit, just MiddleMan approach uses parameter types to distinguish
different methods, while the others use separate sets of methods to distinguish).
We agree that this name conflict problem is not quite popular in practice. But once this issue is
encountered, the effect could be severe, and that existing workarounds might not help a lot. And
since the C++ books, the trait paper as well as other papers mentioned this issue, it might be a
sign that this problem is still an interesting issue in certain cases,
which is worth exploring and taking into consideration.  Finally our work shows that this feature is lightweight, it does not
complicate the semantics a lot and can be easily added to language models.

- "It would strengthen your paper if you would provide an implementation..."

As mentioned in Section 1, we have a prototype implementation available here: 
https://github.com/YanlinWang/MIM/tree/master/Calculus

- "What has been removed from FJ: just downcasts?"

Also constructors and fields.

- "5.2: static invocation confusing..."

* "the method m in J_0 that hierarchically overrides J_1" means, the method m defined like this:
interface J_0 { K m() override J_1 { ... } }
* "e1.J_0@J_1 :: m(e2)" is different from "((J1) e1.m(e2)". The first is a static call while the
second dynamic. The result would be different when the method m gets overridden in an heir of J_0.
* For the question "whether it should be forbidden", we think it is a design choice. If the user
makes  a normal e.m() call, for sure it wouldn't access the super methods. But the static invocation
here gives users the freedom to directly call super methods.

Review C.

- "The proposal seems unnecessarily complex. In NextGen and MixGen ..."

Thanks for the references, we need to add them. The models in these references also show that the
static types can be utilized for method dispatch, which is similar to C++. But the complexity of our
model mainly comes from the feature of hierarchical overriding, which is not introduced by previous
papers. We believe that such complexity can simplify the client code written by programmers.

- "what do you mean by 'reduction model similar to FJ'?"

Sorry for the English, what we mean is that the reduction steps follow the reduction rules in FJ-like
languages, where no static types are tracked.

- "you mention for the first time 'origin'..."

On page 8, "origin" means where the method is defined, we avoid using "original methods" because it
it not introduced yet on page 8. "original method" is a relative concept to hierarchical overriding 
methods.

- "Section 3.2: subtyping is ANTIsymmetric."

Yes, it is not symmetric, we will delete it.

### 2017-12-19 ECOOP18/paper.tex
ECOOP format

\documentclass{llncs}
\pagestyle{plain}

\usepackage{listings}

\usepackage{xspace}
\usepackage{multicol}
\usepackage{microtype}%if unwanted, comment out or use option "draft"
%\usepackage[table,xcdraw]{xcolor} -- clash with preamble.tex
\usepackage{color}
% \usepackage{amsthm}
% \usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{fancyvrb}
\usepackage{url}
%\usepackage{pstricks,pst-node,pst-tree} -- clash with preamble.tex
\usepackage{bbm}
%\usepackage{pgf} -- clash with preamble.tex
\usepackage{multirow}
\usepackage{enumitem}

@@

\let\proof\relax
\let\endproof\relax
\let\mathpar\relax

@@

@@

% \title{Untangling Unintended Conflicts with Hierarchical Dispatch}
% \title{FHJ: A Formal Model for Resolving Unintended Conflicts in Multiple Inheritance}
\title{FHJ: A Formal Model for Hierarchical Dispatching and Overriding}
%\subtitle{\InterfaceBased Programming for the Masses}

% \authorinfo{NoAuthor}
%            {The University of Hong Kong, China}
% %           {\{ylwang,hyzhang,bruno\}@cs.hku.hk}
% %\authorinfo{Marco Servetto}
% %          {Victoria University of Wellington, New Zealand}
% %          {marco.servetto@ecs.vuw.ac.nz}
\author{Yanlin Wang\inst{1} \and Haoyuan Zhang\inst{1} \and Bruno C.d.S. Oliveira\inst{1} \and Marco Servetto\inst{2}}
\institute{The University of Hong Kong \and Victoria University of Wellington \\ \email{\{ylwang,hyzhang,bruno\}@cs.hku.hk, marco.servetto@ecs.vuw.ac.nz}}

@@

\vspace{-4ex}

@@

\begin{comment}
\section*{Acknowledgments}\label{sec:Acknowledgments}

Authors would like to thank YYYYY.
\end{comment}

% \newpage
% \clearpage
% \bibliographystyle{abbrvnat}

\newpage
\bibliographystyle{splncs03}
\bibliography{paper}

\newpage
\appendix
\input{sections/Appendix.tex}
% \input{sections/Appendix_Formalization.tex}
% \input{sections/Translation.tex}
% \input{sections/Appendix_Translation.tex}

>>>>

\documentclass[a4paper,english]{lipics-v2016}
%This is a template for producing LIPIcs articles. 
%See lipics-manual.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
% for section-numbered lemmas etc., use "numberwithinsect"

\usepackage{mathtools} % For \Coloneqq

\usepackage{microtype}%if unwanted, comment out or use option "draft"
\usepackage{listings}

\usepackage{xspace}
\usepackage{multicol}
\usepackage{microtype}%if unwanted, comment out or use option "draft"
%\usepackage[table,xcdraw]{xcolor} -- clash with preamble.tex
\usepackage{color}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{fancyvrb}
\usepackage{url}
%\usepackage{pstricks,pst-node,pst-tree} -- clash with preamble.tex
\usepackage{bbm}
%\usepackage{pgf} -- clash with preamble.tex
\usepackage{multirow}
\usepackage{enumitem}

@@

% \let\proof\relax
% \let\endproof\relax
\let\mathpar\relax

@@

%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory

\bibliographystyle{plainurl}% the recommended bibstyle

% Author macros::begin %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{FHJ: A Formal Model for Hierarchical Dispatching and Overriding}
%\titlerunning{A Sample LIPIcs Article} %optional, in case that the title is too long; the running title should fit into the top page column

%% Please provide for each author the \author and \affil macro, even when authors have the same affiliation, i.e. for each author there needs to be the  \author and \affil macros

\author[1]{John Q. Open}
\author[2]{Joan R. Access}
\affil[1]{Dummy University Computing Laboratory, Address/City, Country\\
  \texttt{open@dummyuniversity.org}}
\affil[2]{Department of Informatics, Dummy College, Address/City, Country\\
  \texttt{access@dummycollege.org}}
\authorrunning{J.\,Q. Open and J.\,R. Access} %mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et. al.'

\Copyright{John Q. Open and Joan R. Access}%mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

\subjclass{Dummy classification -- please refer to \url{http://www.acm.org/about/class/ccs98-html}}% mandatory: Please choose ACM 1998 classifications from http://www.acm.org/about/class/ccs98-html . E.g., cite as "F.1.1 Models of Computation". 
\keywords{Dummy keyword -- please provide 1--5 keywords}% mandatory: Please provide 1-5 keywords
% Author macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Acces}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{CVIT 2016}
\EventAcronym{CVIT}
\EventYear{2016}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}
% Editor-only macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

@@

%%
%% Bibliography
%%

%% Please use bibtex,

\bibliography{paper}
\appendix
%\input{sections/Appendix.tex}

### 2017-12-19 ECOOP18/sections/Introduction.tex
ECOOP format

In contrast to diamond inheritance, the second case of ambiguity
is \textit{unintentional method conflicts}~\cite{scharli03traits}. In
such a case conflicting 
methods do not actually refer to the same feature. 
In a nominal system, methods can be designed for different
functionality, but happen to have the same names (and signatures).
A simple example of this situation is two \lstinline{draw} methods that
are inherited from a deck of cards and a drawable widget. 
In such context, the two \lstinline{draw} methods have very different meanings, 
but they happen to share the same name.
%%This issue was proposed by the trait paper, so-called
When inheritance is used to compose these methods, a compilation 
error happens due to conflicts. However, unlike the diamond problem,
the conflicting methods have very different meanings and do not share a
common parent. We call such a case \textit{triangle inheritance}, in
analogy to diamond inheritance.

@@

C++ and C\# do allow for two
unintentionally conflicting methods to coexist in a class. C\# allows
this by interface multiple inheritance and explicit method
implementations. But since C\# is a single inheritance language, 
it is only possible to \emph{implement} multiple interfaces (but not
multiple classes). %We will have more detailed discussion on this in Section~\ref{sec:relatedwork}.
C++ accepts triangle inheritance and
resolves the ambiguity by specifying the expected path by
\emph{upcasts}. However, neither the C\# or C++ approaches allow
such conflicting methods to be further overridden. 
% However, C++ has
% limited support for virtual methods with unintentional conflicts, and
% it will often throw errors when composing them. This is again
% unsatisfactory because virtual methods are pervasive in OOP and used 
% for code reuse and extensibility. A problem with the C++ approach is
% that programmers can only use either static or dynamic dispatching separately, but dealing
% with unintentional method conflicts seems to require a combination of
% both. 
Some other workarounds or approaches include delegation and
renaming/exclusion in the trait model. However renaming/exclusion 
can break the subtyping relation between a subclass and its parent.
This is not adequate for the class model commonly used in mainstream 
OOP languages, where the subclass is always expected to be a subtype 
of the parent class.

@@

\textit{Hierarchical overriding} is a novel language
mechanism that allows method overriding to be applied
only to one branch of the class hierarchy. Hierarchical overriding 
adds expressive power that is not available in languages such as 
C++ or C\#. Hierarchical overriding allows overriding to work 
for classes with multiple (conflicting) methods sharing the 
same names and signatures. An example is presented in 
Figure~\ref{fig:main}. In this example, there are 4 classes/interfaces.
Two classes \lstinline|Deck| and \lstinline|Drawable| model a deck of cards
 and a drawable widget, respectively. 
The class \lstinline|SafeDeck| adds functionality to check whether the
deck is empty and to prevent drawing a card from an empty deck. 
The interesting class is \lstinline|DrawableSafeDeck|, which inherits 
from both \lstinline|SafeDeck| and \lstinline|Drawable|. Hierarchical
overriding is used in \lstinline|DrawableSafeDeck| to keep two separate \lstinline|draw|
methods for each parent, but overriding \emph{only} the \lstinline|draw|
method coming from \lstinline|Drawable|, in order to draw a widget
with a deck of cards. Note that hierarchical overriding is denoted in the UML
diagram with the notation
\lstinline|draw()|$\shneg$\lstinline|Drawable|, expressing that the \lstinline|draw| method from \lstinline|Drawable|
is overridden. Although in this example only one of the
\lstinline|draw| methods is overridden (and the other is simply inherited), hierarchical overriding
supports multiple conflicting methods to be independently overridden as well.

To present hierarchical overriding and dispatching, we introduce a
formalized model \MIM{} in Section~\ref{sec:formalization} based on
Featherweight Java~\cite{Igarashi01FJ}, together with theorems and
proofs for type soundness. 
In summary, our contributions are:
\begin{itemize}
	\item \textbf{A formalization of a hierarchical dispatching algorithm} that integrates both the static type and dynamic type for method dispatch, and 
	ensures unambiguity as well as extensibility in the presence
        of unintentional method conflicts.
	\item \textbf{Hierarchical overriding:} a novel notion that allows
          methods to override individual branches of the class hierarchy.
	\item \textbf{\name:} a formalized model based on
          Featherweight Java, supporting the above features. 
          We provide the static and dynamic semantics and prove the
          type soundness of the model.
	\item \textbf{Prototype implementation\footnote{https://github.com/YanlinWang/MIM/tree/master/Calculus}:} a
          simple implementation of \MIM{} interpreter in Scala. The implementation can type-check and run variants
          of all the examples shown in this paper.
\end{itemize}

>>>>

In contrast to diamond inheritance, the second case of ambiguity
is \textit{unintentional method conflicts}~\cite{scharli03traits}. In
this case, conflicting 
methods do not actually refer to the same feature. 
In a nominal system, methods can be designed for different
functionality, but happen to have the same names (and signatures).
A simple example of this situation is two \lstinline{draw} methods that
are inherited from a deck of cards and a drawable widget. 
In such context, the two \lstinline{draw} methods have very different meanings, 
but they happen to share the same name.
%%This issue was proposed by the trait paper, so-called
When inheritance is used to compose these methods, a compilation 
error happens due to conflicts. However, unlike the diamond problem,
the conflicting methods have very different meanings and do not share a
common parent. We call such a case \textit{triangle inheritance}, in
analogy to diamond inheritance.

@@

C++ and C\# do allow for two
unintentionally conflicting methods to coexist in a class. C\# allows
this by interface multiple inheritance and explicit method
implementations. But since C\# is a single inheritance language, 
it is only possible to \emph{implement} multiple interfaces (but not
multiple classes). %We will have more detailed discussion on this in Section~\ref{sec:relatedwork}.
C++ accepts triangle inheritance and
resolves the ambiguity by specifying the expected path by
\emph{upcasts}. However, neither the C\# or C++ approaches allow
such conflicting methods to be further overridden. 
% However, C++ has
% limited support for virtual methods with unintentional conflicts, and
% it will often throw errors when composing them. This is again
% unsatisfactory because virtual methods are pervasive in OOP and used 
% for code reuse and extensibility. A problem with the C++ approach is
% that programmers can only use either static or dynamic dispatching separately, but dealing
% with unintentional method conflicts seems to require a combination of
% both. 
Some other workarounds or approaches include delegation and
renaming/exclusion in the trait model. However, renaming/exclusion 
can break the subtyping relation between a subclass and its parent.
This is not adequate for the class model commonly used in mainstream 
OOP languages, where the subclass is always expected to be a subtype 
of the parent class.

@@

\textit{Hierarchical overriding} is a novel language
mechanism that allows method overriding to be applied
only to one branch of the class hierarchy. Hierarchical overriding 
adds expressive power that is not available in languages such as 
C++ or C\#. Hierarchical overriding allows overriding to work 
for classes with multiple (conflicting) methods sharing the 
same names and signatures. An example is presented in 
Figure~\ref{fig:main}. In this example, there are 4 classes/interfaces.
Two classes \lstinline|Deck| and \lstinline|Drawable| model a deck of cards
 and a drawable widget, respectively. 
The class \lstinline|SafeDeck| adds functionality to check whether the
deck is empty and to prevent drawing a card from an empty deck. 
The interesting class is \lstinline|DrawableSafeDeck|, which inherits 
from both \lstinline|SafeDeck| and \lstinline|Drawable|. Hierarchical
overriding is used in \lstinline|DrawableSafeDeck| to keep two separate \lstinline|draw|
methods for each parent, but overriding \emph{only} the \lstinline|draw|
method coming from \lstinline|Drawable|, in order to draw a widget
with a deck of cards. Note that hierarchical overriding is denoted in the UML
diagram with the notation
\lstinline|draw()|$\uparrow$\lstinline|Drawable|, expressing that the \lstinline|draw| method from \lstinline|Drawable|
is overridden. Although in this example only one of the
\lstinline|draw| methods is overridden (and the other is simply inherited), hierarchical overriding
supports multiple conflicting methods to be independently overridden as well.

To present hierarchical overriding and dispatching, we introduce a
formalized model \MIM{} in Section~\ref{sec:formalization} based on
Featherweight Java~\cite{Igarashi01FJ}, together with theorems and
proofs for type soundness. 
In summary, our contributions are:
\begin{itemize}
	\item \textbf{A formalization of the hierarchical dispatching algorithm} that integrates both the static type and dynamic type for method dispatch, and 
	ensures unambiguity as well as extensibility in the presence
        of unintentional method conflicts.
	\item \textbf{Hierarchical overriding:} a novel notion that allows
          methods to override individual branches of the class hierarchy.
	\item \textbf{\name:} a formalized model based on
          Featherweight Java, supporting the above features. 
          We provide the static and dynamic semantics and prove the
          type soundness of the model.
	\item \textbf{Prototype implementation\footnote{https://github.com/YanlinWang/MIM/tree/master/Calculus}:} a
          simple implementation of \MIM{} interpreter in Scala. The implementation can type-check and run variants
          of all the examples shown in this paper.
\end{itemize}
